---
title: "Nonlinear models in R"
output:
  word_document:
    toc: yes
  html_notebook: default
  pdf_document: default
  html_document:
    fig_caption: yes
    toc: yes
---


```{r}
### Load libraies for plots and data analysis
library(tidyr)
library(ggplot2)
library(sjPlot)
library(sjmisc)
library(sjlabelled)

# load response variables

### Read data for specific gene expression
dat1 <- read.csv(file="ENSG115944.csv" )
head(dat1)

# STR data for locus of interest
dat2 <- read.csv(file="ENS115944STRup2.csv",header=FALSE )
dat2[1:5,1:5]

# bring response and STR information together.
dat3 <- merge(dat1,dat2[,c("V1","V42")],by.x="Gene",by.y="V1")
dat3[1:5,]

#We only want the one locus. Drop all other levels from factor variables.
dat4 <- droplevels(dat3)

### Copy the data for detailed analsis requiring madifcations
dat3temp <- dat3
dat3temp$STR <- as.numeric(as.character(dat3temp$V42))

dat3temp[seq(1,nrow(dat3temp),2),"side"] <- "L"
dat3temp[seq(2,nrow(dat3temp),2),"side"] <- "R"
```

*In the next section we will have the following analysis*

The STR(s) are on two independent chanels. To utilize the values to determine significant effects the chanels need to be combined. The following need to be demntrated:

* The chanel has no effect (i.e. left or right), this shows the values are assigned at random.
* The value of the STR should be equal and opposite across the chanel, the values -4 and 4 cancel for example
* The STR and STR^2 are significant along with the interaction of STR and STR^2

```{r}

## Model to determine if there is any significance by side , there should be none.
sidelm <- lm(Response~side,data=dat3temp)
summary(sidelm)

## Model to determine if there is any significance by STR by itself , there should be none.
STRlm <- lm(Response~STR,data=dat3temp)
summary(STRlm)

## Model to determine if there is any significance by STR by itself , there should be none.
STRsidelm <- lm(Response~STR+side,data=dat3temp)
summary(STRsidelm)
tab_model(sidelm,STRlm,STRsidelm)


p1 <- ggplot(dat3temp, aes(as.factor(STR),Response)) + facet_grid(~side) + geom_boxplot()


p1

```

The STR(s) chanels are combined "summed" across.

* The chanel has no effect (i.e. left or right), this shows the values are assigned at random.
* The value of the STR should be equal and opposite across the chanel, the values -4 and 4 cancel for example
* The STR and STR^2 are significant along with the interaction of STR and STR^2


```{r}


### V42 is the start postion we want. start = 42625365

#### Want the STR value to numeric not factor
dat4$STR <- as.numeric(as.character(dat4$V42))
dat4$STRquad <- dat4$STR**2

### Add L/R side can be aribtrary becuase of the model steps above.
### i.e no signifcance when choosing side or STR just need to constent
dat4[seq(1,nrow(dat4),2),"side"] <- "L"
dat4[seq(2,nrow(dat4),2),"side"] <- "R"

dat4[seq(1,nrow(dat4),2),"opside"] <- "R"
dat4[seq(2,nrow(dat4),2),"opside"] <- "L"

### Model with sqaured term.
seplm4a <- lm(dat4$Response ~ dat4$STR + dat4$STRquad + dat4$side )

## STR matter and the squre matters , side L/R does not.
summary(seplm4a)
BIC(seplm4a)

seplm4b <- lm(dat4$Response ~ dat4$STR)
summary(seplm4b)
BIC(seplm4b)



summary(lm(dat4$Response ~ dat4$STR + abs(dat4$STR)))

#dat4$V42 <- factor(dat4$V42,levels(dat4$V42)[order(abs(dat4$STR))])
p2a <- ggplot(dat4, aes(paste0(as.factor(STR),side),Response)) + geom_point() + facet_grid(~opside,scales="free_x")

p2a

summary(lm(Response~paste0(as.factor(STRquad),side),data = dat4))

p2 <- ggplot(dat4, aes(as.factor(STRquad),Response)) + geom_boxplot() + facet_grid(~STR+side,scales="free_x")

p2

```

```{r}

dat5 <- aggregate(dat4[,c("STR","STRquad")],by=list(dat4$Gene,dat4$Response),FUN=sum,na.rm=TRUE)

names(dat5) <- c("Gene","y","STR","STRsqr")


agglm <- lm(y ~ STR+STR*STRsqr,dat=dat5)
summary(agglm)

SSTotal <- var( dat5$y ) * (nrow(dat5)-1)
    SSE     <- sum( agglm$resid^2 )
SSreg   <- SSTotal - SSE

SSTotal
SSreg
SSE


BIC(agglm)

sum(resid(agglm)^2)

mf <- model.frame(agglm)
plot(mf[,"y"],predict(agglm))
xvals <- seq(-24,14,length.out=246)
fitvals <- 0.1458 -0.02701*xvals - 0.001724*xvals**2 + 0.00008863*(xvals*xvals**2)


plot(mf[,"y"],predict(agglm),col=as.factor(mf[,"STR"]),main="Actual v predicted with model fit estimate")
lines(mf[,"y"],fitvals)

plot(mf[,"y"],fitvals,main="Actual v model fit estimate")

#Coefficients:
#              Estimate Std. Error t value Pr(>|t|)    
#(Intercept)  1.458e-01  4.533e-02   3.216 0.001475 ** 
#STR         -2.701e-02  7.974e-03  -3.387 0.000826 ***
#STRsqr      -1.724e-03  5.999e-04  -2.874 0.004416 ** 
#STR:STRsqr   8.863e-05  4.182e-05   2.119 0.035090 * 


p3 <- ggplot(dat5, aes(as.factor(STRsqr),y)) + geom_boxplot() + facet_grid(~STR,scales="free_x")

p3


#agglmcat <- lm(dat5$Group.2~ as.factor(dat5$STR)+as.factor(dat5$STRquad))
#summary(agglmcat)

summary(lm(dat5$y ~ dat5$STR))

dat5$linsum <- dat5$STR + dat5$STRsqr
dat5$linsumint <- dat5$STR + dat5$STRsqr + (dat5$STR*dat5$STRsqr)

linmodsqr <- lm(dat5$y ~ dat5$linsum)

BIC(linmodsqr)

#(Intercept)  0.1240251  0.0434871   2.852 0.004717 ** 
#dat5$linsum -0.0014102  0.0003685  -3.827 0.000165 ***

xvals <- seq(0,300,1)

plot(dat5$linsum,dat5$y)
lines(xvals,0.124-0.00141*xvals)

linmodsqrint <- lm(dat5$y ~ dat5$linsumint)

BIC(linmodsqrint)

summary(linmodsqrint)

xvals <- seq(-6650,1600,5)

plot(dat5$linsumint,dat5$y)
lines(x=0.0000427*xvals)

mf <- model.frame(agglm)
plot(mf[,"y"],predict(agglm))
xvals <- seq(-24,14,length.out=246)
fitvals <- 0.158 -0.0144*xvals - 0.00234*xvals**2
plot(xvals,predict(agglm),col=as.factor(mf[,"STR"]))
lines(xvals,fitvals)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
