---
title: "Nonlinear models in R"
output:
  word_document:
    toc: yes
  html_notebook: default
  pdf_document: default
  html_document:
    fig_caption: yes
    toc: yes
---


```{r}
### Load libraies for plots and data analysis
library(tidyr)
library(ggplot2)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(MASS)
library(tidyverse)
library(caret)
library(leaps)
library(glmnet)
# load response variables

### Read data for specific gene expression
dat1 <- read.csv(file="ENSG115944.csv" )

# STR data for locus of interest
dat2 <- read.csv(file="ENS115944STRup2.csv",header=FALSE )


dat3 <- merge(dat1,dat2,by.x="Gene",by.y="V1")


#We only want the one locus. Drop all other levels from factor variables.
dat4 <- droplevels(dat3)


varit <- 3:67
numcols <- ncol(dat4)-2
resultsout <- NULL
# bring response and STR information together.
for (i in varit) {
#### Want the STR value to numeric not factor
dat4[,i] <- as.numeric(as.character(dat4[,i]))
dat4[,i+numcols] <- dat4[,i]**2

}


dat5 <- aggregate(dat4[,3:ncol(dat4)],by=list(dat4$Gene,dat4$Response),FUN=sum,na.rm=TRUE)

names(dat5)[1:2] <- c("Gene","y")

dat6 <- dat5

varit <- 3:67
numcols <- ncol(dat5)-2
# bring response and STR information together.
for (i in varit) {
#### Want the STR value to numeric not factor
dat6[,i+numcols] <- dat6[,i]*dat6[,i+65]

}


agglm <- lm(y ~ .,dat=dat5[,2:ncol(dat5)])

agglm2 <- lm(y ~ V30+V31+V47+V96+V114,dat=dat5[,2:ncol(dat5)])

agglmf <- lm(y ~ V30+V31*V96+V47*V114,dat=dat5[,2:ncol(dat5)])





SSTotal <- var( dat5$y ) * (nrow(dat5)-1)

    SSE     <- sum( agglm2$resid^2 )
SSreg   <- SSTotal - SSE
BICreg <- BIC(agglm2)

  SSEf     <- sum( agglmf$resid^2 )
SSregf   <- SSTotal - SSEf
BICregf <- BIC(agglmf)


SSTotal

SSE
SSEf
  
SSreg  
SSregf

BICreg
BICregf

#resultsout <- rbind(resultsout,cbind(i,BICreg,BICregf,SSTotal,SSreg,SSE,SSregf,SSEf))

```

```{r}

full.model <- lm(y ~ .,dat=dat6[,2:ncol(dat6)])
# Stepwise regression model
step.model <- stepAIC(full.model, direction = "both", 
                      trace = FALSE)

BIC(step.model)

thestep <- lm(formula = y ~ V4 + V18 + V20 + V30 + V31 + V38 + V48 + V51 + 
    V59 + V73 + V83 + V84 + V88 + V95 + V96 + V97 + V101 + V109 + 
    V117 + V118 + V125 + V135 + V153 + V154 + V159 + V162 + V169 + 
    V174 + V179 + V190 + V142, data = dat6[, 2:ncol(dat6)])

summary(thestep)

  SSEstep     <- sum( thestep$resid^2 )
SSregstep   <- SSTotal - SSEstep
BICregstep <- BIC(thestep)



thestep1 <- lm(formula = y ~ V4 + V18 + V20 + V30 + V38 + V48 +  
    V59 + V73 + V83 + V84 + V88 + V95 + V96 + V101 + V109 + 
    V117 + V135 + V153 + V159 + V169 + 
    V174 + V179 + V142, data = dat6[, 2:ncol(dat6)])

plot(predict(thestep1),dat6[,"y"],main=paste0("SS:",round(SSTotal,2)," REG: ",round(SSregstep,2)," ER: ",round(SSEstep,2)," BIC: ",round(BICregstep,2)))

SSEstep1     <- sum( thestep1$resid^2)
SSregstep1   <- SSTotal - SSEstep1
BICregstep1 <- BIC(thestep1)


xvals <- seq(-24,24,length.out=246)

plot(xvals,dat6[,"y"],type="p",col="green",main=paste0("SS:",round(SSTotal,2)," REG: ",round(SSregstep1,2)," ER: ",round(SSEstep1,2)," BIC: ",round(BICregstep1,2)))
lines(xvals,predict(thestep1),type="p",col="red")

 

SSTotal
SSEstep
SSEstep1
SSregstep
SSregstep1
BICregstep
BICregstep1
```

So what’s the bottom line? In general, it might be best to use AIC and BIC together in model selection. For example, in selecting the number of latent classes in a model, if BIC points to a three-class model and AIC points to a five-class model, it makes sense to select from models with 3, 4 and 5 latent classes. AIC is better in situations when a false negative finding would be considered more misleading than a false positive, and BIC is better in situations where a false positive is as misleading as, or more misleading than, a false negative.


```{r eval=FALSE, echo=FALSE}

dat7 <- dat5

varit <- 3:67
numcols <- ncol(dat5)-2
# bring response and STR information together.
for (i in varit) {

#### Want the STR value to numeric not factor
dat7[,numcols:(numcols+(66-i))] <- dat5[,i]*dat5[,i:67]
numcols <- ncol(dat7)
}


training.samples <- dat7$y %>% 
  createDataPartition(p = 0.8, list = FALSE)
train.data  <- dat7[training.samples, ]
test.data <- dat7[-training.samples, ]


cv.lasso <- cv.glmnet(as.matrix(train.data[,3:ncol(train.data)]), train.data$y, alpha = 1, family = "gaussian")

lasso1<-glmnet(x=as.matrix(train.data[,3:ncol(train.data)]),train.data$y,family="gaussian",alpha=1,lambda = cv.lasso$lamdba.1se)


x.test <- model.matrix(y ~., test.data)[,-1]

probabilities <- lasso1 %>% predict(newx = x.test)


observed.classes <- test.data$y


lasso1 %>% predict(test.data, type = "response")


full.model2 <- lm(y ~ .,dat=dat7[,2:ncol(dat7)])
# Stepwise regression model
step.model2 <- stepAIC(full.model2, direction = "both", 
                      trace = FALSE)

BIC(step.model2)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
